{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or: How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-09-20 thesis/related.tex
more rela

Testing approaches for transactional programs are sparsely represented in the literature so far.
Although several related works \cite{tm-correctness,tm-completeness,specifying-verifying-tm}
are building up to formal proofs of the correctness of underlying TM {\em implementations},
Landslide is the first I know of to verify client programs thereof.
McRT STM \cite{mc-tm-with-spin} uses SPIN \cite{spin} to model check an STM implementation
up to 2 threads running 1 transaction each with up to 3 memory accesses.
This kind of verification,
analogous to \sect{\ref{sec:education-pebbles-tests}}'s {\tt mutex\_test},
is an important stepping stone for trusting the results Landslide will provide.
STAMP \cite{stamp} is a benchmark suite transactional programs,
implemented using the OpenTM interface \cite{opentm},
used by many papers in the field to evaluate the performance of both STM and HTM implementations alike.
Even so, the more recent Stampede suite \cite{scalable-tm} argues that STAMP's benchmarks
were constructed under a programming model poorly-suited to fully take advantage of HTM's performance,
and that scalable HTM programs should
seek to minimize incidental conflicts and to handle aborts more flexibly than with blind retry loops.
The programming complexity needed to achieve these goals calls,
of course,
for correspondingly advanced verification approaches such as Landslide.
Finally, TxRace \cite{txrace}
tests non-transactional programs for data races
by inserting HTM calls via compiler instrumentation,
% they don't actually seem to check the failure reason lmao
% i'll phrase this in a slightly weasel way to point out that's what they *should* be doing
% and that in turn motivates the stuff i did
% even tho they use a diff way to filter them post hoc (well, they'd still need to, bc of false sharing)
relying on conflict aborts to point out access pairs that would be unsafe in the original program.
This citation perhaps half belongs in \sect{\ref{sec:related-data-race}} as well;
I include it here to highlight the importance of Landslide's
ability to distinguish different abort reasons (\sect{\ref{sec:txn-abort-codes}}).

@@

% .. rather than just executing it a bunch of times and seeing what behaviours fall out ..

@@

>>>>

Testing approaches for transactional programs are sparsely represented in the literature so far.
Although several related works \cite{tm-correctness,tm-completeness,specifying-verifying-tm}
are building up to formal proofs of the correctness of underlying TM {\em implementations},
Landslide is the first I know of to verify client programs thereof.
McRT STM \cite{mc-tm-with-spin} uses SPIN \cite{spin} to model check an STM implementation
up to 2 threads running 1 transaction each with up to 3 memory accesses.
This kind of verification,
analogous to \sect{\ref{sec:education-pebbles-tests}}'s {\tt mutex\_test},
is an important stepping stone for trusting the results Landslide will provide.
STAMP \cite{stamp} is a benchmark suite transactional programs,
implemented using the OpenTM interface \cite{opentm},
used by many papers in the field to evaluate the performance of both STM and HTM implementations alike,
although as discussed in \sect{\ref{sec:tm-eval-exp-setup}},
focuses more on performance than on interesting concurrency properties.
% stamp bugfixes - https://github.com/mfs409/transmem/blob/master/benchmarks/stamp_c/VERSIONS.
% stamp need not be considered harmful put here
Even so, the more recent Stampede suite \cite{scalable-tm} argues that STAMP's benchmarks
were constructed under a programming model poorly-suited to fully take advantage of HTM's performance,
and that scalable HTM programs should
seek to minimize incidental conflicts and to handle aborts more flexibly than with blind retry loops.
The programming complexity needed to achieve these goals calls,
of course,
for correspondingly advanced verification approaches such as Landslide.
Finally, TxRace \cite{txrace}
tests non-transactional programs for data races
by inserting HTM calls via compiler instrumentation,
% they don't actually seem to check the failure reason lmao
% i'll phrase this in a slightly weasel way to point out that's what they *should* be doing
% and that in turn motivates the stuff i did
% even tho they use a diff way to filter them post hoc (well, they'd still need to, bc of false sharing)
relying on conflict aborts to point out access pairs that would be unsafe in the original program.
This citation perhaps half belongs in \sect{\ref{sec:related-data-race}} as well;
I include it here to highlight the importance of Landslide's
ability to distinguish different abort reasons (\sect{\ref{sec:txn-abort-codes}}).

@@

It should come as no surprise that many avenues of research towards writing correct programs
have been explored
apart from just executing them a bunch of times to check all the thread interleavings.
Though not as directly related to Landslide as the works referenced above,
this section explores such other approaches,
ranging from building safety guarantees into a language's type system
to checking, proving, and/or enforcing execution properties post-hoc.

@@

\subsubsection{Symbolic execution}

Analogous to SCT,
which seeks good coverage of possible thread execution paths under schedule nondeterminism,
another popular testing approach is symbolic execution \cite{symbolic-execution},
which seeks good coverage of possible flow control paths under input nondeterminism.
Symbolic executors
abstract a program's variables and use constraint solving engines such as Z3 \cite{z3}
to work backwards and, effectively,
synthesize combinations of test inputs which can lead to a failure.
KLEE \cite{klee}, one well-known and open-source implementation of this technique,
offers over 90\% code coverage on average across many tests,
often outdoing that achieved by programmers' own hand-written tests.
Later, Contessa \cite{contessa} extended symbolic execution to include concurrency nondeterminism as well,
by using a DPOR-like analysis on individual execution traces
then incorporating reordering possibilities into its SMT constraints.
%rather than using explicit scheduling as SCT does.
This simultaneously exercises both input and schedule nondeterminism,
and achieves good bug-finding performance in practice,
but does not provide the same verification guarantees as repeated DPOR iterations with explicit scheduling.
Exploring both kinds of state space at once thoroughly enough to provide strong verification
is undoubtedly subject to further state space explosion,
and remains future work.
